### 2022-09-16 paper_new/200-base-language.tex
More work on paper

% !TEX root = document.tex

\chapter{\label{chap:baselang}Base Language}

The base language that was implemented is the Calculus of Constructions \autocite{Coquand_Huet_1988}, with a syntax somewhat similar to that of Haskell. One extra feature was added that is not present in the Calculus of Constructions, that is, let bindings.

\section{Syntax}

The syntax of the base language is defined in SDF3, the syntax definition language of Spoofax. The definition is very similar to that of a simply typed lambda calculus, except that types and expressions are a single sort.

\begin{lstlisting}
context-free sorts
	Expr

context-free syntax
	Expr.Let = [let [ID] = [Expr]; [Expr]]
	Expr.Type = "Type"
	Expr.Var = ID
	Expr.FnType = ID ":" Expr "->" Expr {right}
	Expr.FnConstruct = "\\" ID ":" Expr "." Expr
	Expr.FnDestruct = Expr Expr {left}
	Expr = "(" Expr ")" {bracket}

context-free priorities
	Expr.Type > Expr.Var > Expr.FnType > Expr.FnDestruct 
		> Expr.FnConstruct > Expr.Let
\end{lstlisting}

\section{Static Analysis}

\section{How scope graphs are used}

To type-check the base language, we need to use Statix, and scope graphs. This section describes how scope graphs are used.

The scope graph only has a single type of edge, we call them \verb|P| edges (parent edges). It also only has a single relation, called \verb|name|. This name stores a \verb|NameEntry|, which can be either a \verb|NameType|, which stores the type of a name, or a \verb|NameSubst|, which stores a substitution corresponding to a name.

\begin{lstlisting}
signature sorts
	NameEntry
constructors
	NameType : Expr -> NameEntry
	NameSubst : scope * Expr -> NameEntry
relations
	name : ID -> NameEntry
name-resolution	labels P
\end{lstlisting}

These are all the definitions we will need to type-check programs. We will create a name with \verb|NameType| as entry for function parameters, and we will use \verb|NameSubst| for let bindings and function application in beta reduction.

\section{Garbage}

- Base language is calculus of constructions with lets + type assertions (move type assertions to sec 5?)
- Describe syntax of the base language?
- Describe rules of base language (statix syntax or mathy?)
- Scope graphs for substitutions + scopes