### 2022-09-29 paper/000-abstract.tex
Work on writing

% !TEX root = document.tex

When working with the Spoofax workbench, the Statix meta-language can be used for the specification of Static Semantics. To provide these advantages to as many language developers as possible, Statix aims to cover a broad range of languages and type-systems. However, no attempts have been made to express dependently typed languages in Statix. Dependently typed languages are unique in that types may depend on values. Type-checking them comes with unique challenges, such as the need to evaluate terms during type-checking. We present how to make a dependent language by implementing the calculus of constructions in Statix. Next, we discuss the difficulties that come from using names, rather than De Bruijn indices. Finally, we explore features such as type Inference and inductive data types.