### 2022-11-14 paper-thesis/chapters/31-comparing-with-lambdapi.tex

\chapter{A comparison with implementations in logical frameworks}
\label{ch:comp-lambdapi}