### 2022-11-29 paper-thesis/chapters/02-background-spoofax.tex
Work on agda.lp

% !TEX root = document.tex
\chapter{\label{chap:bg-spoofax}Background: Spoofax}