### 2022-12-07 paper-thesis/chapters/02-background-spoofax.tex
Background section for spoofax

% !TEX root = document.tex
\chapter{\label{chap:bg-spoofax}Background: Spoofax}

The Spoofax Language Workbench \cite{spoofax} is a platform to develop domain-specific and
general-purpose programming languages. Each aspect of the language is defined in one of three metalanguages, each with their own purpose. From these languages Spoofax generates a parser, type checker and other useful tools.

The metalanguage SDF3 is used to define the syntax of the language \cite{sdf3}, which is then parsed using Spoofax's SGLR parser \cite{sdf3_parser}. The parser produces a abstract syntax tree in the ATerm format \cite{aterm}. A pretty printer is also generated based on this syntax definition.

Next, the metalanguage Statix \cite{scopes_as_types} is used to define the static semantics of the language. It uses scope graphs to represent binding and typing information of the language. How this works is discussed in section \ref{sec:statix}.

Finally, Stratego \cite{stratego} is used to define program transformations, using rewrite rules. This can be used for modifications to the AST, but also to build transformations to a different language (such as compiling to assembly).

\section{\label{sec:statix} Statix and Scope graphs}

In Statix, scope graphs are used to represent a programs name and type information \cite{scopes_as_types}. A scope graph is a graph that consists of the following components:

\begin{itemize}
	\item \textit{Scopes} represent "a region in a program that behaves uniformly with respect to name resolution". These scopes are modeled as nodes in the graph. Its graph representation is shown in \autoref{fig:graph-notation-scope}.
	\item \textit{Labeled, directed edges} model visibility relations between scopes. For example, $\scopet{1} \scopeedget[\lblt{P}] \scopet{2}$ indicates that the graph contains an edge from \scopet{1} to \scopet{2} with label \lblt{P}. Its graph representation can be seen in \autoref{fig:graph-notation-edge}.
	\item \textit{Declarations} model a datum (a piece of information, like the type of a variable) under a relation symbol in a scope. Its textual notation is $\scopet{1} \typeedget[rel] d$, meaning datum $d$ is declared in scope $\scopet{1}$ under relation \textit{rel}. Its pictorial equivalent is shown in \autoref{fig:graph-notation-data}.
\end{itemize}

\begin{figure}[!h]
	\begin{framed}
		\begin{subfigure}{0.2\linewidth}
			\centering
			\begin{tikzpicture}[scopegraph, node distance = 5em and 3.0em]
				\node[scope] (s1) {$1$};
			\end{tikzpicture}
			\caption{Scope}
			\label{fig:graph-notation-scope}
		\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
			\centering
			\begin{tikzpicture}[scopegraph, node distance = 5em and 3.0em]
				\node[scope] (s1) {$1$};
				\node[scope, right of = s1] (s2) {$2$};
				\draw (s1) edge[lbl={P}] (s2);
			\end{tikzpicture}
			\caption{Edge}
			\label{fig:graph-notation-edge}
		\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
			\centering
			\begin{tikzpicture}[scopegraph, node distance = 5em and 3.0em]
				\node[scope] (s1) {$1$};
				\node[type, right of = s1] (s2) {$\texttt{x : INT()}$};
				\draw (s1) edge[type, lbl={rel}] (s2);
			\end{tikzpicture}
			\caption{Data}
			\label{fig:graph-notation-data}
		\end{subfigure}
	\end{framed}
	\caption{Scope Graph notation}
\end{figure}

Information can be retrieved from scope graphs using \textit{queries}. A query from a scope traverses the scope graph to find datums that match certain conditions. The result of a query is a list of (path, datum) tuples. A query has several parameters:
\begin{itemize}
	\item The relation to query. Only datums declared under this relation will be returned by the query.
	\item Path well-formedness condition: a regular expression of labels that specifies which paths are well-formed. Only datums that are reached through a path whose edge labels are in the language described by the regular expression are included in the query result.
	\item A match predicate, taking a single datum as input. Only datums that satisfy this predicate will be included in the query result. Its default value is \verb|true|, meaning that all datums under the specified relation that satisfy the path well-formedness condition are returned.
	\item Result comparison parameters:
	\begin{itemize}
		\item A strict partial order on paths, described by less-than relations on labels. This relation defines a prefix-order on paths.
		\item A `shadow' predicate, which takes two datums as input. Its default value is \verb|false|, meaning that no shadowing is performed.
	\end{itemize}
	When the result set contains multiple datums, and it holds for a datum $d$ that there exists another datum $d'$ which path is strictly smaller that $d$, and the shadow predicate holds, then $d$ is removed from the result set.
\end{itemize}

% CREDIT FOR THIS EXPLANATION TO ARON ZWAAN
% HOW TO CREDIT?