### 2023-01-18 paper-thesis/chapters/12-editor-services.tex
Add the semantic code completion chapter

% !TEX root = document.tex
\chapter{\label{chap:editor-services}Semantic Code Completion}

We explored how semantic code completion presented by Pelsmaeker et al.~\cite{codecompletion} applies to dependently typed languages. Code completion is an editor service in IDEs that proposes code fragments for the user to insert at the caret position in their code.

\section{Setup required}
To set up editor services, we followed the steps at \todo{cite} \url{https://spoofax.dev/spoofax-pie/develop/guide/static-semantics/code-completion/}. To be precise, we followed the following steps:
\begin{enumerate}
	\item Add \verb|tego-runtime {}| and \verb|code-completion {}| to the \verb|spoofax.cfg| file, to enable code completion.
	\item Add the following rules to the \verb|main.str2| file, to pre-process and post-process the AST for code completion.
	\begin{lstlisting}
rules
	downgrade-placeholders = downgrade-placeholders-MyLang
	upgrade-placeholders   = upgrade-placeholders-MyLang
	is-inj                 = is-MyLang-inj-cons
	pp-partial             = pp-partial-MyLang-string
	pre-analyze            = explicate-injections-MyLang
	post-analyze           = implicate-injections-MyLang
	\end{lstlisting}
	\item For each rule define a predicate that accepts a placeholder where a syntactic sort is permitted. For our language, those are the following:
	\begin{lstlisting}
expectBetaEq((_, Expr-Plhdr()), _).
expectBetaEq(_, (_, Expr-Plhdr())).
betaReduceHead((_, Expr-Plhdr())) = _.
betaReduce((_, Expr-Plhdr())) = _.
typeOfExpr(_, Expr-Plhdr()) = _.
programOk(Start-Plhdr()).
	\end{lstlisting}

\end{enumerate}

\section{How well it works}
In order to get suggestions, we can now insert a placeholder \verb|[[Expr]]| and press \verb|ctrl + space| in the editor to get semantic suggestions.

It works well, only showing completions that are semantically relevant. For example, given the following code it only suggests expressions that can be booleans:
\begin{lstlisting}
let f = \b: Bool. b;
f [[Expr]]
\end{lstlisting}
This would return the following suggestions: (Note that \verb|f| and \verb|Type| are not suggested, since they cannot have type \verb|Bool|)
 \begin{itemize}
 	\item \verb|[[Expr]] [[Expr]]|
 	\item \verb|let [[ID]] = [[Expr]]; [[Expr]]|
 	\item \verb|true|
 	\item \verb|false|
 	\item \verb|if [[Expr]] then [[Expr]] else [[Expr]] end|
 \end{itemize}