### 2023-01-18 paper-thesis/chapters/40-discussing-statix.tex
Meeting results

\chapter{The usability of Spoofax for defining dependently typed languages}