### 2023-01-25 paper-thesis/chapters/40-discussing-statix.tex
Work from yesterday

\chapter{Ergonomics of Spoofax}

Now that we've implemented a feature-rich dependently typed language in Statix, this chapter discusses the experience.

\section{Statix}

- expressive language
- more control over metavariables
- nested constraints (datatypes reduce)
- generic functions? reverse/concat for lists
- debugging (statix tests are not great)

\section{Spoofax}

- errors
- performance
- testing works well
-