### 2023-02-03 paper-thesis/chapters/02-background-spoofax.tex
Start reworking spoofax chapter

% !TEX root = document.tex
\chapter{\label{chap:bg-spoofax}Background: Spoofax}

This chapter will explain the concepts behind the Spoofax Language Workbench. Readers already familiar with Spoofax are recommended to only skim through this chapter.

The Spoofax Language Workbench~\cite{spoofax} is a platform to develop domain-specific and
general-purpose programming languages. Each aspect of the language is defined in one the meta-languages, each with their own purpose. From these specifications Spoofax generates a parser, type checker and other useful tools. The meta-languages that are used in this thesis are discussed in the sections below:

\section{SDF3}

The meta-language SDF3 is used to define the syntax of the language~\cite{sdf3}, which is then parsed using Spoofax's SGLR parser~\cite{sdf3_parser}. The parser produces a abstract syntax tree in the ATerm format~\cite{aterm}. A pretty printer is also generated based on this syntax definition.

\paragraph{Syntax definitions}
The syntax of

\paragraph{ATerm format}

\section[Statix and Scope graphs]{\label{sec:statix}Statix and Scope graphs\protect\footnote{Note: This section is adapted from "Composable Type System Specification using Heterogeneous Scope Graphs".~\cite[sect. 4.1.2]{Zwaan21}}}

In Statix, scope graphs are used to represent a programs name and type information~\cite{scopes_as_types}~\cite{nameres}. A scope graph is a graph that consists of the following components:

\begin{itemize}
	\item \textit{Scopes} represent "a region in a program that behaves uniformly with respect to name resolution". These scopes are modeled as nodes in the graph. Its graph representation is shown in \autoref{fig:graph-notation-scope}.
	\item \textit{Labeled, directed edges} model visibility relations between scopes. For example, $\scopet{1} \scopeedget[\lblt{P}] \scopet{2}$ indicates that the graph contains an edge from \scopet{1} to \scopet{2} with label \lblt{P}. Its graph representation can be seen in \autoref{fig:graph-notation-edge}.
	\item \textit{Declarations} model a datum (a piece of information, like the type of a variable) under a relation symbol in a scope. Its textual notation is $\scopet{1} \typeedget[rel] d$, meaning datum $d$ is declared in scope $\scopet{1}$ under relation \textit{rel}. Its pictorial equivalent is shown in \autoref{fig:graph-notation-data}.
\end{itemize}

\begin{figure}[!h]
	\begin{framed}
		\begin{subfigure}{0.2\linewidth}
			\centering
			\begin{tikzpicture}[scopegraph, node distance = 5em and 3.0em]
				\node[scope] (s1) {$1$};
			\end{tikzpicture}
			\caption{Scope}
			\label{fig:graph-notation-scope}
		\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
			\centering
			\begin{tikzpicture}[scopegraph, node distance = 5em and 3.0em]
				\node[scope] (s1) {$1$};
				\node[scope, right of = s1] (s2) {$2$};
				\draw (s1) edge[lbl={P}] (s2);
			\end{tikzpicture}
			\caption{Edge}
			\label{fig:graph-notation-edge}
		\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
			\centering
			\begin{tikzpicture}[scopegraph, node distance = 5em and 3.0em]
				\node[scope] (s1) {$1$};
				\node[type, right of = s1] (s2) {$\texttt{x : INT()}$};
				\draw (s1) edge[type, lbl={rel}] (s2);
			\end{tikzpicture}
			\caption{Data}
			\label{fig:graph-notation-data}
		\end{subfigure}
	\end{framed}
	\caption{Scope Graph notation}
\end{figure}

Information can be retrieved from scope graphs using \textit{queries}. A query from a scope traverses the scope graph to find datums that match certain conditions. The result of a query is a list of (path, datum) tuples. A query has several parameters:
\begin{itemize}
	\item The relation to query. Only datums declared under this relation will be returned by the query.
	\item Path well-formedness condition: a regular expression of labels that specifies which paths are well-formed. Only datums that are reached through a path whose edge labels are in the language described by the regular expression are included in the query result.
	\item A match predicate, taking a single datum as input. Only datums that satisfy this predicate will be included in the query result. Its default value is \verb|true|, meaning that all datums under the specified relation that satisfy the path well-formedness condition are returned.
\end{itemize}