### 2023-02-07 paper-thesis/chapters/03-background-dependent-types.tex
Improve chapter 3

% !TEX root = document.tex
\chapter{\label{chap:bg-dp}Background: Dependent Types}

This chapter will explain the concepts behind dependent types. Readers already familiar with Dependent Types are recommended to only skim through this chapter.

\section{What are Dependent Types?}

A dependent type system is a type system in which types may depend on values. This means it is possible to have a type such as \verb|Vec n|, which is a vector of exactly length \verb|n|. Note that the value of \verb|n| does not have to be known at compile time, it is more powerful than that.

The biggest advantage of dependent types is that they increase the expressiveness of a type system a lot, for example:
\begin{lstlisting}
append : (A: Set) -> (n : Nat) -> Vec A n -> Vec A n -> Vec A (n + n)
\end{lstlisting}
A \verb|Vec A n| is a list with elements of type \verb|A| that is exactly length \verb|n| (an integer, which is a value!). This append function appends two \verb|Vecs| of equal length, returning a \verb|Vec| of length \verb|n + n|. This is an example of a type-level computation.

Dependent types are useful because dependent type systems allow for the Curry-Howard Correspondence~\cite{chc}. Using this correspondence, types correspond to propositions and a term of a type correspond to a proof of a proposition. Thus if our type system is sound\footnote{The type system we will define in chapter \ref{chap:baselang} is not sound, we will fix this in chapter \ref{ch:universes}.}, that is, we cannot create a term of an empty type, then we can use our type systems to prove things.

\section{The Calculus of Constructions}

The lambda cube is a framework to categorize programming languages based on whether terms and types can depend on each other~\cite{lambda_cube}. Specifically, it categorizes languages on three axes: \footnote{Note that the "Terms can depend on terms" axis is missing, this is because a language where terms cannot depend on terms cannot compute anything and is thus pretty useless.}

\begin{enumerate}
	\item Terms can depend on types. This corresponds to type-polymorphic (generic) functions. For example, it allows us to define the polymorphic identity function:
	\begin{lstlisting}
id : (T : Type) -> T -> T
	\end{lstlisting}
	\item Types can depend on types. This corresponds to type-polymorphic (generic) datatypes. For example, it allows to the define the \verb|List T| type, a list with items of type T.

	\item Types can depend on terms. This corresponds to dependent types.
\end{enumerate}

The \textit{Calculus of Constructions} is the language that has all three of these features, it will be implemented in chapter \ref{chap:baselang}.