### 2023-02-28 paper-thesis/chapters/24-backend.tex
Start backend

\chapter{Back-end}
\label{ch:backend}

We have now fully built the front-end (parser and type checker) of our language. However, the language also needs a back-end to be usable.