### 2023-03-08 paper-thesis/chapters/24-backend.tex

\chapter{Back-end}
\label{ch:backend}

We have now fully built the front-end (parser and type checker) of our language. However, the language also needs a back-end to be usable. A back-end can be either an interpreter or a compiler, we will explore both options in this chapter.

\section{Interpreter}
It is valuable to define an interpreter for our language, because it is a way to run a program without relying on the definition of the language you are compiling to. For a dependently typed language specifically it is easy to add on, as we already needed to define the dynamic semantics of our language for type checking anyways.

However, this big advantage is also a disadvantage: Dynamic semantics would usually be defined in Stratego. We need to find a way to re-use the definition from Statix, since we don't want to repeat the definition of the dynamic semantics.

\subsection{Implementation}

Luckily, Stratego has a strategy built in that runs a Statix predicate on a term: \verb|stx-evaluate|. Using this strategy, the actual implementation of our interpreter in Stratego is just one line: Match an expression \verb|e| and run the functional predicate \verb|interpret| from the \verb|main.stx| file on it.

\begin{lstlisting}
interpret : e -> <stx-evaluate(|"main", "main!interpret")> [e]
\end{lstlisting}

Then, in the \verb|main.stx| file we define \verb|interpret| to be beta reducing the expression in an empty scope.

\begin{lstlisting}
interpret(e) = betaReduce((sEmpty(), e)).
\end{lstlisting}

These two lines are the only required changes to implement an interpreter\footnote{Other than changing the project config to define a button in the Spoofax UI that when pressed runs the interpret strategy on a given term}. This shows that it is very straight-forward to implement an interpreter.

\subsection{Discussion of defining in Statix versus Stratego}

An advantage of implementing the dynamic semantics in Statix instead of Stratego is that we can use scope graphs as a part of our dynamic semantics rules.

A disadvantage is that Stratego is specifically optimized for these transformations~\cite{fcpmc}. Therefore, the resulting interpreter will be slower when implemented in Statix. \todo{comparison}

\section{Compiler}
The alternative to writing an interpreter is writing a compiler to another language. In order to explore this avenue, we wrote a compiler to Clojure, a dynamically typed dialect of Lisp that runs on the JVM~\cite{clojure}. The advantage of a compiling to such a high-level language is that the actual compilation steps are relatively straight-forward, but it still shows that we can write a compiler for a dependently typed language in Spoofax.

Both languages support first-class functions, so we can map most constructs of our language directly onto constructs of Clojure. Therefore, most of our compiler has rules that look like the example below, where the constructors are compiled recursively using the \verb|compile-expr| strategy, ignoring the specified types. The constructors starting with \verb|C| are constructors from Clojure. 
\begin{lstlisting}
compile-expr : FnConstruct(x, _, b) -> 
  CFn(<compile-id> x, <compile-expr> b)
\end{lstlisting}

Types require a bit more attention. Since our language has no way to match on types, types can actually be completely eliminated during type checking. The easiest way to do this is to compile all types to an arbitrary value (we chose the integer 0), since the value is irrelevant anyways. A more efficient solution could be to implement type erasure~\cite[Section 23.6]{tapl}, where type-level functions are completely removed. We decided against this as the goal of this section is to show that we can compile the language with ease, not to write an efficient compiler.

-analysis is hard, agda doesn't do it
- if compiled language doesn't matter
- erasure annotations (idris/agda)

Finally, inductive data types are a bit harder to compile, as Clojure does not have an equivalent to sum types on which we can match. \todo{finish this after implementing}
- agda to scheme compiler
- atoms voor constructors